% !TEX TS-program = XeLaTeX
% !TEX encoding = UTF-8 Unicode

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%	大连理工大学硕士论文 XeLaTeX 模版 —— 封面文件 cover.tex
%	版本：0.6
%	最后更新：2010.11.15
%	修改者：Yuri (E-mail: yuri_1985@163.com)
%	编译环境：Ubuntu 10.04 + TeXLive 2010 + TeXworks
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cdegree{硕~~士~~学~~位~~论~~文}
\ctitle{基于独立成分分析的煤仓沉降数据分析}
\etitle{Data Analysis of the Settlement of Coal bulker based on Independent Component Analysis}

% 根据需要添加字符间距
\csubject{{\quad\;}大地测量学与测量工程}
\cauthor{{\quad\;}苏~运~强}
\cauthorno{{\quad\;}470920341}
\csupervisor{{\quad\;}赵~~国~~忱　教授}

% 这里默认使用最后编译的时间，也可自行给定日期，注意汉字和数字之间的空格。
\cdate{{\quad\;}\the\year~年~\the\month~月~\the\day~日}

\cabstract{
随着社会的发展，人类对能源的需求越来越大。当前我国的主要一次能源仍然是煤炭。
更加规范的管理煤炭的生产、运输、使用等在对于安全生产、保护环境等是非常重要
的。其中，煤仓是煤炭的生产、运输、使用等环节中都非常重要的设施，其可以作为各
环节之间的缓冲，也可以避免煤炭露天堆放产生的消耗和环境污染，因此在煤矿、煤炭
销售、热电厂、煤化工厂等中都得到了广泛的使用。

煤仓是一种大型的工业建筑，其直径甚至可以达到百余米，高可达数十米，装
载能力可以达到数十万吨。如此巨大的建筑、如此多的载荷将是对设计、施工和运营的
巨大考验，也是对地基和煤仓壁的巨大考验。

为了确保煤仓的安全运营，同时为以后的设计提供原始资料等
必须要对煤仓进行定期的沉降观测，并采用一些数学方法处理对数据进行处理，
分析煤仓的沉降规律，预测煤仓的进一步下沉。

当前有多种技术应用到了建筑物的变形监测中，对于不是特别大的建筑物的沉降，
水准方法仍然是最合适、最常用的方法。

对于采集到的数据需要进行处理：传统的变形几何分析主要包括参考点的稳定性分析、
观测值的平差处理和质量评定以及变形模型参数估计等内容；
为了考虑变形体在不同状态之间具有的时间关联性，20世纪70年代以后，
陆续引入了的时间序列分析方法、基于数字信号处理的数字滤波技术分离时效分量、
变形的卡尔曼滤波模型等方法。

在对观测数据进行处理得同时，人们也试图对变形进行物理解释。
产生了主要的3类方法：为统计分析法、确定函数法和混合模型法。

首先编写了程序来对煤仓沉降数据进行处理，生成相应的报表、绘制沉降曲线，
极大提高了工作效率。

本文试图将当前主要用于图像特征提取、语音分离等的独立成分分析(ICA)技术应用
煤仓的沉降数据处理。
将煤仓上多个不同点沉降观测序列作为多个信号，对其应用独立成分分析技术，
具体使用FastICA算法，以及先使用主成分分析(PCA)获取相应的独立成分，
并与煤仓的装载量、气温、时间等进行对比，
使用相关系数、互信息以及灰关联系数作为指标分析它们之间的关系。
取得了一定的成果。
}

\ckeywords{煤仓；沉降观测；变形监测数据处理；独立成分分析；FastICA}

\eabstract{
With the development of society, more energy is being required. 
At present, coal is still the primary energy for China.
More standardized management of coal production, transportation, use, etc. 
is very important for the safe production and environmental protection. 
Of. Among them, coal bunkers are widely used in coal production, 
transportation and other sectors and are very important facilities, 
which can be used as the buffer between sections, 
and can also used to avoid the environmental pollution generated by open dumping of coal, 
so the coal mines, coal Sales, power plant, coal chemical plants, etc.

Coal bunker is a large industrial building, 
or even up to hundred meters in diameter, up to tens of meters tall.
Carrying capacity can reach several hundred thousand tons. 
Such a huge building, so much the load will be a great challenge to 
the design, construction and operation,
of course also to the foundation and the great test of bunker wall.

In order to ensure the safe operation of the coal bunker, 
and advising for the future design, etc.
the settlement observation on the coal bunker must be regular, 
and using some of the mathematical approach for data processing,
to analyze coal bunker settlement rules and to predict the further sinking coal bunker.

There are many techniques currently applied to the 
deformation monitoring of the building. For buildings that are not to giant,
the leveling method is still the most appropriate and commonly used method.

The data collected needs to be processed: 
traditional geometric deformation analysis includes
analysis of the stability of the reference point,
Observations of the adjustment processing and quality assessment, 
and the deformation parameter estimation, etc.;
in order to consider the deformation of the body in the time 
between the different states of relevance, 
since the 1970s, gradually introduced in time series analysis method, 
based on digital signal processing digital filtering technology component separation time,
deformation of the Kalman filter model and other methods.

At the same time with trying best to the processing of observational data, 
people are also trying to make the physical interpretation of the deformation.
And produced three main types of methods: the statistical analysis, 
to determine the function and mixed model method.

Here, first, wrote a program to process the data: 
generate reports, drawing settlement curve, etc.
Thus work efficiency was greatly improved.

This article attempts use Independent Component Analysis (ICA) 
which is mainly used on image feature extraction, 
speech separation, for coal bunker settlement data processing.
Will settlement observation sequences of several different points on the coal bunker be multiple signals, 
and apply them to independent component analysis to got several independent components.
In practice, FastICA algorithm can be used either directly or be used with principal component analysis (PCA). 
With obtain the corresponding independent component, the loading, temperature and time were compared.
With using the Pearson correlation coefficient, mutual information, 
and gray relational coefficients as indicators of the relationship between them,
some results achieved.
}

\ekeywords{Coal bulker; Settlement observation; Deformation monitoring; Independent Component Analysis; FastICA}

\makecover 
